% \documentclass[12pt,a4paper]{article}

\documentclass[lic, english, shortabstract]{iithesis}

\makeatletter
% Define \@pdfcolorstack as a count if it's missing (needed for pdftex)
\@ifundefined{@pdfcolorstack}{\csname newcount\endcsname\@pdfcolorstack}{}

% Fix section numbering to not include chapter (0.)
\renewcommand{\thesection}{\arabic{section}}
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\makeatother

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{lmodern}
\usepackage{microtype} 
\usepackage{csquotes}

\usepackage[backend=biber, style=ieee]{biblatex}
\addbibresource{bibliography.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{geometry}
\geometry{
    a4paper,
    left=25mm, 
    right=25mm,
    top=25mm,
    bottom=25mm,
    headheight=15pt,
    headsep=10mm,
    footskip=12mm,
    bindingoffset=0mm
}
\usepackage{setspace}
\onehalfspacing 

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.5pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsmath, amssymb, amsthm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{xcolor}
\usepackage{graphicx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{tcolorbox}
\tcbuselibrary{skins, breakable, theorems}
\usepackage{float}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{titlesec}
\titlespacing*{\section}{0pt}{14pt plus 4pt minus 2pt}{10pt plus 2pt minus 2pt}
\titlespacing*{\subsection}{0pt}{12pt plus 3pt minus 2pt}{8pt plus 2pt minus 2pt} 
\usepackage{enumitem}
\setlist[enumerate]{itemsep=0pt, topsep=0.25\baselineskip, parsep=0pt, partopsep=0pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=black!70!blue,
    filecolor=black!70!red,      
    urlcolor=black!70!blue,
    citecolor=black!70!blue,
    pdftitle={Longest Increasing Subsequence},
    pdfauthor={Mikołaj Owczarz}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\theoremstyle{plain}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\newcommand{\Lis}{\textsc{LIS}}
\newcommand{\LisLen}{\textsc{LIS-LENGTH}}
\newcommand{\Psort}{\textsc{Patience Sorting}}
\newcommand{\Ppass}{\textsc{Patience Pass}}

\newcommand{\problembox}[3]{
    \begin{tcolorbox}[
        colback=gray!5!white,      
        colframe=gray!60!black,    
        arc=3pt,                   
        boxrule=1pt,               
        left=8pt, right=8pt, top=8pt, bottom=8pt
    ]
        \textbf{Problem:} #1 \\[1.5ex]
        \textbf{Input:} #2 \\[1.5ex]
        \textbf{Output:} #3
    \end{tcolorbox}
}



\tcbset{
    statementbox/.style={
        enhanced,
        breakable,          
        boxrule=0.8pt,      
        colframe=black!75,  
        colback=gray!5,     
        arc=3pt,            
        outer arc=3pt,
        left=8pt, right=8pt, top=6pt, bottom=6pt,
        before skip=12pt,   
        after skip=12pt
    }
}

\newtheorem{claim}[theorem]{Claim}
\tcolorboxenvironment{claim}{statementbox}

\newtheorem{observation}[theorem]{Observation}
\tcolorboxenvironment{observation}{statementbox}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\tcolorboxenvironment{definition}{statementbox}

% Pamietaj o madrym (zgodnym z logicznym rozbiorem zdania oraz estetyka) recznym
% zlamaniu wierszy w temacie pracy, zwlaszcza tego w jezyku pracy. Uzyj do tego
% polecenia \fmlinebreak.
\polishtitle    {Problem najdłuższego rosnącego podciągu w małej pamięci}
\englishtitle   {Longest Increasing Subsequence Problem in Small Space, revisited}
\polishabstract {
    Dla danej tablicy wejściowej, naszym celem jest znalezienie najdłuższego podciągu rosnącego lub jego długości, przy wykorzystaniu sparametryzowanej pamięci.
    Powszechnie wiadomo, że problem ten można rozwiązać w czasie $\mathcal{O}(n \log n)$ bez ograniczeń pamięciowych.
    Dla sparametryzowanej pamięci $s$, gdzie $\sqrt{n} \leq s \leq n$, przedstawiamy ulepszoną analizę algorytmu \Psort\ ze sparametryzowaną pamięcią, opisanego w pracy \cite{kiyomi2017space}. 
    Pokazujemy, że obliczanie długości najdłuższego podciągu rosnącego tym przy użyciu tego algorytmu działa w czasie $\mathcal{O}(\max(n \log n, \frac{n^2}{s}))$. 
    Przedstawiamy również ulepszony algorytm typu dziel i rządź dla wyznaczania podciągu, który działa w czasie $\mathcal{O}(\max(n \log n \log \log n, \frac{n^2}{s}))$, wykorzystując przy tym $\mathcal{O}(s \log n)$ bitów.
}
\englishabstract{
    Given an input array of comparable elements, we aim to find either a longest increasing subsequence, or its length using parametrised space.
It is well known that this problem can be solved in $\mathcal{O}(n \log n)$ time without memory constraints. For parametrised memory -- $s$, where
$\sqrt{n} \leq s \leq n$, we present an improved analysis of space-parametrised \Psort\ algorithm introduced in \cite{kiyomi2017space}. We show that the length computation runs in $\mathcal{O}(\max(n \log n, \frac{n^2}{s}))$ time.
We also present an improved divide-and-conquer algorithm for obtaining the sequence, which runs in $\mathcal{O}(\max(n \log n \log \log n, \frac{n^2}{s}))$ time while using $\mathcal{O}(s \log n)$ bits.
}
\author         {Mikołaj Owczarz}
\advisor        {dr Karol Pokorski}
\date          {\today}                     % Data zlozenia pracy
\transcriptnum {345235}                     % Numer indeksu
\advisorgen    {dr. Karola Pokorskiego} % Nazwisko promotora w dopelniaczu

\begin{document}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\section{Introduction[EN]}

The longest increasing subsequence (LIS for short) problem is a classic algorithmic problem. 
The input is a sequence of $n$ numbers $A = (a_1, a_2, \dots, a_n)$, and our goal is either to find the length of the LIS -- 
the largest integer $k$ for which there exist indices $i_1 < i_2 < \dots < i_k$ such that $a_{i_1} < a_{i_2} < \dots < a_{i_k}$, or to extract such subsequence.

\subsection{Previous results}
Efficient algorithms for finding the LIS are well known, for example, the \Psort\ algorithm devised by Mallows \cite{mallows}, which runs in $\mathcal{O}(n \log n)$ time for a sequence of length $n$ while using $\mathcal{O}(n)$ words of working space.
Given a sequence of $n$ elements, \Psort\ assigns each element to a \textit{pile}, where each pile contains information about the longest increasing subsequence of the input that ends in that element. 
While the \Psort\ algorithm achieves the lower bound in terms of time complexity for finding the LIS \cite{lis-lb}, the question of how little memory one needs to efficiently solve the problem remains unanswered. 
A variation of \Psort\ is presented in \cite{kiyomi2017space}, which, for a given memory parameter $s$, can find the length of the LIS in $\mathcal{O}(\frac{n^2}{s} \log n)$ time and compute the actual subsequence in $\mathcal{O}(\frac{n^2}{s} \log^2 n)$ time, but only for $\sqrt{n} \leq s$.

\subsection{Our contribution}
In this paper, we improve the bounds achieved in \cite{kiyomi2017space}. For finding the length of the LIS, we conduct a tighter time analysis, showing that the algorithm proposed in \cite{kiyomi2017space} actually runs in $\mathcal{O}(\max(n \log n, \frac{n^2}{s}))$ time,
while maintaining the $\mathcal{O}(s)$ bound for the number of words in the working memory. 
Next, we observe that an alternative, simpler divide-and-conquer algorithm can be derived, which is more efficient than the one provided in \cite{kiyomi2017space}.
Combined with the tighter analysis for the LIS length computation, it runs in $\mathcal{O}(\max(n \log n \log \log n, \frac{n^2}{s}))$ time.

\subsection{Related work}
The LIS problem, in addition to its classic setting, has many heavily-researched variations.
For example, we could ask for the longest increasing subsequence only on some part of the input by making range queries \cite{range-lis}.
We can also ask to maintain the LIS in a dynamic setting \cite{dynamic-lis, dynamic-lis-lb}. Another interesting question is 
the approximation of the LIS in any of these settings \cite{approx-lis, approx-lcs}. 

\subsection{Applications}
As one of the most classic algorithmic problems, the LIS problem has a variety of applications across diverse scientific domains.
In bioinformatics, the LIS is a fundamental tool for genome rearrangement analysis and sequence alignment, where it is often used to find maximum weight traces or to solve the Longest Common Subsequence (LCS) problem for some special cases \cite{delcher1999alignment}. 
It also plays a crucial role in statistical physics and probability theory; notably, the distribution of the length of the LIS in random permutations is a central topic in the study of random matrices and the Tracy-Widom distribution \cite{baik1999distribution}. 
Additionally, LIS algorithms are used in data analysis for identifying trends in time-series data and in version control systems for sequence differencing, for example in the well-known \texttt{diff} command.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction[PL]}

Problem najdłuższego podciągu rosnącego (w skrócie LIS, od ang. \textit{longest increasing subsequence}) jest klasycznym problemem algorytmicznym. 
Dla danych wejściowych -- sekwencji $n$ liczb $A = (a_1, a_2, \dots, a_n)$ naszym celem jest znalezienie długości LIS -– 
największej liczby całkowitej $k$, dla której istnieją indeksy $i_1 < i_2 < \dots < i_k$ takie, że $a_{i_1} < a_{i_2} < \dots < a_{i_k}$, lub wyznaczenie takiego podciągu.

\subsection{Dotychczasowe wyniki}
Wydajne algorytmy znajdowania LIS są dobrze znane.
Przykładem jest algorytm \Psort\ opracowany przez Mallowsa \cite{mallows}, który działa w czasie $\mathcal{O}(n \log n)$ dla ciągu o długości $n$, wykorzystując przy tym $\mathcal{O}(n)$ słów pamięci roboczej.
Mając dany ciąg $n$ elementów, algorytm \Psort\ przypisuje każdy element do pewnego \textit{stosu} (\textit{pile}), gdzie każdy stos zawiera informacje o najdłuższym rosnącym podciągu wejścia, który kończy się tym elementem. 
Chociaż algorytm \Psort\ osiąga dolną granicę złożoności czasowej dla problemu znajdowania LIS \cite{lis-lb}, pytanie o to, jak mało pamięci potrzeba do jego wydajnego rozwiązania, pozostaje bez odpowiedzi. 
W pracy \cite{kiyomi2017space} zaprezentowano wariant algorytmu \Psort, który dla zadanego parametru pamięci $s$ potrafi znaleźć długość LIS w czasie $\mathcal{O}(\frac{n^2}{s} \log n)$ oraz wyznaczyć sam podciąg w czasie $\mathcal{O}(\frac{n^2}{s} \log^2 n)$, jednakże z dodatkowym warunkiem -- $\sqrt{n} \leq s$.

\subsection{Nasz wkład}
W niniejszej pracy poprawiamy ograniczenia uzyskane w \cite{kiyomi2017space}.
W przypadku obliczania długości LIS przeprowadzamy ściślejszą analizę czasową, pokazując, że algorytm zaproponowany w \cite{kiyomi2017space} w rzeczywistości działa w czasie $\mathcal{O}(\max(n \log n, \frac{n^2}{s}))$,
zachowując jednocześnie ograniczenie $\mathcal{O}(s)$ na liczbę słów w pamięci roboczej. 
Następnie zauważamy, że można wyprowadzić alternatywny, prostszy algorytm typu „dziel i rządź”, który jest wydajniejszy niż ten przedstawiony w \cite{kiyomi2017space}
i w połączeniu ze ściślejszą analizą obliczania długości LIS, działa on w czasie $\mathcal{O}(\max(n \log n \log \log n, \frac{n^2}{s}))$.

\subsection{Prace powiązane}
Problem LIS, oprócz swojego klasycznego sformułowania, posiada wiele intensywnie badanych wariantów.
Na przykład, możemy szukać najdłuższego podciągu rosnącego tylko na pewnej części wejścia, wykonując zapytania na spójnych przedziałach ciągu wejściowego \cite{range-lis}.
Możemy również rozważać dynamiczne utrzymanie LIS \cite{dynamic-lis, dynamic-lis-lb}. Innym interesującym zagadnieniem jest 
aproksymacja LIS w dowolnym z tych wariantów \cite{approx-lis, approx-lcs}.

\subsection{Zastosowania}
Jako jeden z najbardziej klasycznych problemów algorytmicznych, problem LIS znajduje szereg zastosowań w różnych dziedzinach nauki.
W bioinformatyce LIS jest podstawowym narzędziem do analizy rearanżacji genomów oraz dopasowywania sekwencji (ang. \textit{sequence alignment}), gdzie często służy do znajdowania ścieżek o maksymalnej wadze (ang. \textit{maximum weight traces}) lub do rozwiązywania problemu najdłuższego wspólnego podciągu (LCS) dla pewnych szczególnych przypadków \cite{delcher1999alignment}. 
Odgrywa również kluczową rolę w fizyce statystycznej i teorii prawdopodobieństwa; w szczególności rozkład długości LIS w losowych permutacjach jest głównym tematem badań nad macierzami losowymi i rozkładem Tracy'ego-Widoma \cite{baik1999distribution}. 
Ponadto algorytmy LIS są wykorzystywane w analizie danych do identyfikacji trendów w szeregach czasowych, a także w systemach kontroli wersji do wyszukiwania różnic w sekwencjach, czego przykładem jest dobrze znane polecenie \texttt{diff}.

\section{Preliminaries}

We define two variants of the Longest Increasing Subsequence problem.

\problembox
{Longest increasing Subsequence (\Lis)}
{A sequence of $n$ numbers $A = (a_1, a_2, \dots, a_n)$.}
{The longest subsequence $(a_{i_1}, a_{i_2}, \dots, a_{i_k})$ of $A$ such that $i_1 < i_2 < \dots < i_k$ and $a_{i_1} < a_{i_2} < \dots < a_{i_k}$.}

\problembox
{Length of \Lis \ (\LisLen)}
{A sequence of $n$ numbers $A = (a_1, a_2, \dots, a_n)$.}
{$k$ -- the largest integer such that there exist an increasing subsequence in $A$ of length $k$}


We use the RAM model, where the input sequence is read-only, and the working memory is limited to $\mathcal{O}(s)$ words.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Patience Sorting}

Our algorithm is based on the classic \Psort. Let us define basic notions used throughout the algorithm. The idea of the algorithm is to maintain \textit{piles} -- stacks of integers such that for all $x \in A$, $x$ belongs to the $i$-th pile if and only if the \textit{longest increasing subsequence} of $A$ ending in $x$ has length $i$.
Since both classic and space-parametrised variants of the \Psort\ algorithm are based on sequential passes through the input array, we define
pile operations with respect to the time when they are being used.

\begin{definition}
For integers $i, t$, we refer to the state of the $i$-th pile after reading the first $t$ elements of the input sequence as $P_{i}^{t}$, and $P_i$ will represent the $i$-th pile after reading the whole input sequence.
\end{definition}
Now we define some useful operations that can be conducted on piles.
Let $P$ be some state of the pile (for example $P_3^4$ or $P_2$), then:
\begin{itemize}
\item $|P|$ is the number of elements that belong to $P$
\item $\operatorname{top}(P)$ refers to the element on the top of the pile (the one that was added most recently to it), or null if $P$ is empty.
\item For an integer $j$, $P[j]$ is the $j$-th element of $P$, where $\operatorname{top}(P) = P[|P|]$, and $P[1]$ is the element on the bottom of $P$. 
\end{itemize}

It follows from the pile definition that if we are solving the \LisLen\ problem, then our goal is to compute the largest integer $k$ such that $P_k$ is non-empty.
The classic \Psort\ algorithm achieves that by constructing all piles during a single pass through the input sequence. When reading the $t$-th element of $A$ -- $a_t$, we find
the smallest $i$ such that $\operatorname{top}(P_i^{t-1}) \geq a_t$ and we push $a_t$ to the top of the $i$-th pile, or we create a new, rightmost pile with a single element $a_t$.
By definition, each element belongs to exactly one pile. Thus, for every $t$, if $a_t$ is in the $i$-th pile then $P_i^t$ is $P_i^{t-1}$ with $a_t$ pushed onto it, and $P_j^t = P_j^{t-1}$ for all $j \neq i$.
Since copying and storing the pile states at every step would introduce time and memory issues, and the state transitions are straightforward, we can naturally construct the piles on the fly while iterating through $A$.
Thus, we omit the upper indices in the implementation of \Psort\ algorithm below.

\begin{algorithm}
    \caption{\Psort}
    \label{alg:psort}
    \begin{algorithmic}[1]
        \State initialize $P_{0}$ with the single element $-\infty$
        \State set $k := 0$ \Comment{\textcolor{gray}{$k$ will be the index of the rightmost non-empty pile}}
        \For{$a_t \in A$}
            \If{$a_t > \text{top} (P_{k})$}
                \State create an empty pile -- $P_{k+1}$, and push $a_t$ to $P_{k+1}$
                \State set $k := k + 1$
            \Else
                \State set $h$ to be the smallest index with $x \leq \operatorname{top}(P_h)$, and push $x$ to $P_h$
            \EndIf
        \EndFor
        \State \Return $k$
  \end{algorithmic}
\end{algorithm}


\begin{observation}\label{obs:pile}(\cite{burnstein})
Algorithm~\ref{alg:psort} correctly computes piles, in other words: \\
Element $a_t \in P_i$ if and only if a \textit{longest increasing subsequence} of $A$ ending in $a_t$ is of length $i$
\end{observation}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Space-parametrised Patience Sorting}
Now we define the space-parametrised variant of \Psort\ that originates from \cite{kiyomi2017space}.

Assume that our working memory is limited to $\mathcal{O}(s)$ words, and $\sqrt{n} \leq s \leq n$.
Similarly to the classic \Psort\ we will process the input array sequentially while maintaining piles. However, since our memory is limited and we cannot afford
to keep all piles in the memory, we will pass through the input multiple times. Each time we will store only one, chosen pile and top elements of some other piles. The idea is clearly presented in \cite{kiyomi2017space} (Lemma 4.2, Lemma 4.3).

More precisely, now we define a procedure \Ppass\ that, given the $i$-th pile $P_i$, performs two passes through the input array.
During the first pass it computes an index $j$ such that $i + s \leq j \leq i + 2s \land |P_j| \leq s$ or $j$ is the length of a \Lis. 
During the second pass, now with given $P_i$ and $j$ it will compute whole $P_j$, or only the $\operatorname{top}(P_j)$ if $j$ is the length of a \Lis.

\begin{observation}\label{obs:filtering} (\cite{kiyomi2017space}, Observation 4.1)
Suppose we are reading $a_t$, and $a_t \in P_j$. Let $i$ be a fixed integer. \\
If $j \neq i$, then:
\begin{itemize}
\item $a_t < \operatorname{top}(P_i^{t-1})$ if $j < i$
\item $a_t > \operatorname{top}(P_i^{t-1})$ if $j > i$
\end{itemize} 
\end{observation}

Recall that in the \Psort\ algorithm, for an element $a_t$, to determine to which pile it belongs, it suffices to know only the values of the top elements of the piles at step $t-1$. 
Moreover, by Observation~\ref{obs:filtering}, if we want to check whether this index -- call it $k$, belongs to a specific interval---in our case, whether $i \leq k \leq i + 2s$---it suffices to know only the top elements of the piles whose indices are also in this interval.
Below, we define the \textit{filtering} method for the \Ppass\ procedure.
Suppose we are processing $a_t$. We can encounter one of the following three cases: 
\begin{enumerate}
\item If $a_t = \operatorname{top}(P_i^t)$, then $k = i$.
\item Otherwise, if $a_t < \operatorname{top}(P_i^{t-1})$, then, by Observation~\ref{obs:filtering}, $k < i$, and we ignore $a_t$.
\item Otherwise, $a_t > \operatorname{top}(P_i^{t-1})$, and again, by Observation~\ref{obs:filtering}, $k > i$.
Let $r := \operatorname{top}(P_{i + 2s}^{t-1})$. Then:
    \begin{enumerate}
    \item If $r$ exists and $a_t > r$, then $k > i + 2s$, and we ignore $a_t$.
    \item Otherwise, if $r$ is undefined or $a_t \leq r$, we have $i < k \leq i + 2s$, and $k$ is the smallest index such that $a_t \leq \operatorname{top}(P_k^{t-1})$.
    \end{enumerate}
\end{enumerate}
Observe that if the first case is false, then $\operatorname{top}(P_i^{t-1}) = \operatorname{top}(P_i^{t})$; thus, exactly one of these cases holds.

Notice that if we have the whole $P_i$ in memory, we can process the first two cases in constant time by maintaining a pointer to the most recently read element $h = \operatorname{top}(P_i^{t-1})$ in $P_i$.
In the first case, we compare $a_t$ to the element in $P_i$ that is after $h$. We can determine whether we are in the second or third case by comparing $a_t$ and $h$.

It is easy to see that if we maintain $\operatorname{top}(P_k^{t-1})$ for all $i < k \leq i + 2s$, then we can distinguish cases 3(a) and 3(b) in constant time and find the value of $k$ in 3(b) in $\mathcal{O}(\log s)$ time.
We will call the search in 3(b) an \textit{expensive filter}. Below, we present the \Ppass\ implementation (Algorithm 3 in \cite{kiyomi2017space}).

\begin{algorithm}
    \caption{First pass in \Ppass, assume $P_i$ is given, and $|P_i| < s$}
    \begin{algorithmic}[1]
    \State set $h := 0$ \Comment{\textcolor{gray}{$h$ points to the most recently read element in $P_{i}$}}
    \State set $r := i$ \Comment{\textcolor{gray}{Is the index of the current rightmost pile}}
    \State set $p_{i+1}, p_{i+2}, \ldots, p_{i+2s}$ to $\infty$ \Comment{\textcolor{gray}{$p_k$ is the current $\operatorname{top}(P_{k})$}}
    \State set $c_{i+1}, c_{i+2}, \ldots, c_{i+2s}$ to $0$ \Comment{\textcolor{gray}{$c_k$ is the current size of $P_k$}}
    \For{$ x \in A$} 

        \If{$x$ is $P_i[h+1]$} \Comment{\textcolor{gray}{$x$ is the next element in $P_i$}}
            \State set $h := h + 1$ 
            \State continue
        \ElsIf {$x < P_i[h]$ or $x > p_{i+2s}$} \Comment{\textcolor{gray}{$x$ belongs to the $P_k$ where $k < i \vee k > i+2s$}}
            \State continue
        \EndIf

        \If{$x > p_r$}
            \State set $r := r + 1$ and $k := r$
        \Else
            \State set $k$ to be the smallest such that $x < p_k$
        \EndIf
            
        \State set $p_k := x$ and $c_k := c_k + 1$
    \EndFor
    
    \State set $j$ to be the biggest, such that $0 < c_j \leq s$
    
    \Return $j$
    \end{algorithmic}
\end{algorithm}

After this pass, if $r < i + 2s$, we can conclude that $r$ is the length of the \Lis, since no element belongs to a pile with a larger index.
Then, we can set $j := r$ and return $j$ as the \LisLen, along with $p_j = \operatorname{top}(P_j)$. 
It is easy to see how to modify the procedure to compute the whole $P_j$ if $j$ is given: we can maintain an additional stack and push the elements that belong to the $j$-th pile onto it.

\begin{observation} \label{obs:ppass-advance}
Since $s > \sqrt{n}$ there are at most $\frac{n}{s} \leq s$ piles with more than $s$ elements which implies that $j > i + s$.
\end{observation}

\begin{observation} \label{obs:ppass-time}
The time complexity of a single \Ppass\ is $\mathcal{O}(n + W \log s )$, where $W$ is the number of calls to the expensive filter during the procedure.
\end{observation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{New results}

The complexity analysis for the space-parametrised \Psort\ in \cite{kiyomi2017space} for finding \LisLen\ yielded an $\mathcal{O}(\frac{n^2}{s} \log n)$ time.
By providing a tighter estimation of the number of calls to the \textit{expensive filter} procedure, we prove that this algorithm actually runs in $\mathcal{O}(\max(n \log n, \frac{n^2}{s}))$ time.
Furthermore, \cite{kiyomi2017space} presented a divide-and-conquer algorithm for computing a \Lis\ in $\mathcal{O}(\frac{n^2}{s} \log^2 n)$ time.
We provide a simpler and more efficient divide-and-conquer strategy that, combined with the improved bound for computing the \LisLen, runs in $\mathcal{O}(\max(n \log n \log \log n, \frac{n^2}{s}))$ time.


\subsection{Length computation}
We provide a tighter time complexity analysis of the \Psort\ algorithm presented in \cite{kiyomi2017space}. 
We show that the algorithm runs in $\mathcal{O}(\max(n \log n, \frac{n^2}{s}))$ time, improving upon the previous $\mathcal{O}(\frac{n^2}{s} \log n)$ bound.

\begin{definition}
    Let $j_i$ be the index of the pile kept in memory after $i$-th \Ppass\ and we assume $j_0=0$.
\end{definition}

\begin{definition}
Let $W_i$ be the number of \textit{expensive filter} operations during the $i$-th \Ppass.
\end{definition}

\begin{definition}
Let $K$ be the total number of \Ppass -es.
\end{definition}

By using Observation~\ref{obs:ppass-time} we can see that time complexity of space-parametrised \Psort\ for $\sqrt{n} \leq s \leq n$ is:
\[
T(n,s) \in \mathcal{O}(nK + \log s \sum_{i=1}^K W_i).
\]

\begin{claim} \label{claim:linear-filters}
The total number of expensive filter operations is linear, i.e.,
\[
\sum_{i=1}^K W_i \in \mathcal{O}(n).
\]
\end{claim}

\begin{proof}
For each pass $i \in \{1, \dots, K\}$, let us define the active interval $I_i = (j_{i-1}, j_{i-1}+2s]$.

Fix a pass $i$. During the first phase of this pass, the \textit{expensive filter} is invoked only for elements whose pile index belongs to $I_i$. Once the new boundary $j_i$ is determined, the second phase invokes the \textit{expensive filter} only if the pile index is in $(j_{i-1}, j_i]$. 
By \Ppass\ construction, we know that $j_i \leq j_{i-1} + 2s$, which means $(j_{i-1}, j_i] \subseteq I_i$. Thus, we can see that:
\[
W_i \le \sum_{k \in I_i} |P_k| + \sum_{k=j_{i-1}+1}^{j_i} |P_k| \leq 2\sum_{k \in I_i} |P_k|.
\]
This gives us: 
\[
\sum_{i=1}^K W_i \le 2\sum_{i=1}^K \sum_{k \in I_i} |P_k|.
\]

By \Ppass\ construction (Observation~\ref{obs:ppass-advance}) we have $j_i \ge j_{i-1}+s$  or $j_i$ is the length of \Lis.
This implies that the left endpoints of the intervals $I_i$ advance by at least $s$ for all but the last pass.
Observe that when not counting the last pass each interval $I_i$ has a width of exactly $2s$, any pile index $k$ can belong to at most two intervals $I$.
Therefore, summing the pile sizes over all intervals yields at most twice the total number of elements:
\[
\sum_{i=1}^{K-1} \sum_{k \in I_i} |P_k| + \sum_{k \in I_K} |P_k| \le 2\sum_{k=1}^n |P_k| + \sum_{k=1}^n |P_k|  = 3n.
\]
Thus, we get $\sum_{i=1}^K W_i \le 6n \in \mathcal{O}(n)$.
\end{proof}

Because after each \Ppass\ we advance at least by $s$ piles and the largest possible pile index is $n$, $K \in \mathcal{O}(\frac{n}{s})$, which gives us the total running time:
\[
T(n,s) \leq cn \frac{n}{s} + c'n\log s
\]
For some constants $c, c'$. It is easy to see that $T(n, s) \in \mathcal{O}(\max(n\log n, \frac{n^2}{s}))$ as desired.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{LIS computation}
To compute the \Lis, we present an alternative divide-and-conquer algorithm, which combined with the improved length bound, runs in $\mathcal{O}(\max(n \log n \log \log n, \frac{n^2}{s}))$ time. This improves on the earlier $\mathcal{O}(\frac{n^2}{s} \log^2 n)$ result.

\begin{definition}
For $x \in A$ let $\textsc{Lis}_x$ be the \textit{longest increasing subsequence} of $A$ ending in $x$.
\end{definition}

\begin{definition}
For an integer $x$ and sequence of numbers $A$, let $A_{\leq x}$ be the subsequence of $A$ consisting of elements that are equal of smaller than $x$. Analogously we define $A_{>x}$.
\begin{itemize}
\item $L_{\leq x}$ is a \Lis\ for $A_{\leq x}$.
\item $L_{>x}$ is a \Lis\ for $A_{>x}$.
\end{itemize}
\end{definition}

Let $A_{\ell}$ and $A_r$ be the left and right halves of $A$ respectively. Our goal is to efficiently find an element $m$ such that
concatenation of $L_{\leq m}$ for $A_\ell$ and $L_{> m}$ for $A_r$ is a \Lis\ for original input $A$. 
Having such element would allow us to recursively solve \Lis\ for $A_\ell$ and $A_r$ independently. We will show that such element can be found during a single \LisLen\ computation without changing its time and memory complexity.

\begin{definition}
For $x \in A$, let $\operatorname{pred}(x) \in A_\ell$ be such that there is a $\textsc{Lis}_x$ in $A$
such that the last element of it belonging to $A_\ell$ is pred$(x)$, or null if the length of $\textsc{Lis}_x$ is 1 and $x \notin A_\ell$.
\end{definition}

\begin{claim} Let $a_t \in P_i$, then \\
pred$(a_t) = \begin{cases}
a_t & \text{if } a_t \in A_\ell \\
\text{null} & \text{if } a_t \in A_r \land i = 1 \\
\text{pred}(y) & \text{where } y = \operatorname{top}(P_{i-1}^{t-1})
\end{cases}$
\end{claim}

\begin{proof}
Case when $a_t \in A_\ell$ is trivial. Assume $a_t \in A_r$. Now, if $a_t \in P_1$, then the $\textsc{Lis}_{a_t}$ is of
length $1$, thus by definition $\operatorname{pred}(a_t) = \text{null}$. Now, assume that $a_t \in A_r \land a_t \in P_i \land i > 1$. Then  
$\textsc{Lis}_{a_t}$ is of length $i$. Since $y = \text{top}(P_{i-1}^{t-1})$ is smaller than $a_t$ (by Observation~\ref{obs:filtering}), and $a_t$ is later in $A$ than $y$, if we append $a_t$ to $\textsc{Lis}_y$ it is still a valid $increasing\ subsequence$.
Notice that since length of $\textsc{Lis}_y$ is $i-1$, by appending $a_t$ to it we obtain $\textsc{Lis}_{a_t}$. Since $a_t \in A_r$ the last element belonging to $A_\ell$ in $\textsc{Lis}_{a_t}$ is the same one that in $\textsc{Lis}_y$ making $\operatorname{pred}(a_t) = \operatorname{pred}(y)$.  
\end{proof}

\begin{observation}
Let $a \in A$ be an element that belongs to the last pile. 
Then, for $m = \operatorname{pred}(a)$ concatenation of $L_{\leq m}$ for $A_\ell$ and $L_{> m}$ for $A_r$ is a \Lis.
\end{observation}

\begin{proof}
This follows directly from the definition of $\operatorname{pred}$. Let us choose an element $a$ from the last pile. 
Consider a $\textsc{Lis}_a$ where $m$ is the last element from $A_\ell$ belonging to this sequence.
Since $a$ is located in the last pile, $\textsc{Lis}_a$ is a \Lis.
Because $\textsc{Lis}_a$ is an increasing sequence and $m$ is in it, there exist increasing subsequences $L_1 \text{ and } L_2$, such that $\textsc{Lis}_a$ is their concatenation and $m$ is the last element of $L_1$.
It immediately follows that each element of $L_1$ is less than or equal to $m$ and each element of $L_2$ is greater than $m$.
Since $m \in A_\ell$, $L_1$ is a subsequence of $A_\ell$. Because $m$ is the last element from $A_\ell$ in $\textsc{Lis}_a$, $L_2$ is a subsequence of $A_r$.
Therefore, since $\textsc{Lis}_a$ is a \Lis, $L_1$ is $L_{\leq m}$ and $L_2$ is $L_{> m}$.
\end{proof}

It can be easily seen how to modify the \Ppass\ procedure and store the pred of an element using only $\mathcal{O}(n)$ time per \Ppass.
We can also see that to compute pred for any element that is not filtered out during a \Ppass\ we can use pred values 
and memoising the additional value only for the $\mathcal{O}(s)$ elements that we already store in memory, thus keeping both time and memory complexity with respect to big-$\mathcal{O}$ notation.
We deduce that the previously mentioned element $m$ can be $\operatorname{pred}(x)$ for any $x$ in the last pile and since computing pred values does not change the time and memory complexity of the \Ppass\ 

\begin{observation}The time complexity for solving the \Lis\ can be written as:
$$T(n, s) = \begin{cases}
c'n \log n \text{ if } n \leq s \\
2T(\frac{n}{2}, s) + c\operatorname{max}(n \log n, \frac{n^2}{s}) \text{ otherwise}
\end{cases}$$
For some constants $c, c'$, which is $\mathcal{O}(\operatorname{max}(n \log n \log \log n, \frac{n^2}{s}))$.
\end{observation}
This is because we can find the element $m$ during a single \Ppass\ which runs in $\mathcal{O}(\operatorname{max}(n \log n, \frac{n^2}{s}))$ time and then solve on two halves of the input independently.

To justify the complexity bound suppose that we claim $T(n, s) = 2T(\frac{n}{2}, s) + c_1n \log n + c_2\frac{n^2}{s}$ when $n > s$ for some constants $c_1, c_2$.
Notice that $n \log n$ dominates over $\frac{n^2}{s}$ when $s > \frac{n}{\log n}$, and there are at most $\log \log n$ such levels until the base case $n \leq s$ is reached, contributing at most $\mathcal{O}(n \log n \log \log n)$ to the total work.
Otherwise, when $\frac{n^2}{s}$ dominates, the recursion yields $\mathcal{O}(\frac{n^2}{s})$. 
Therefore, the final time complexity is $\mathcal{O}(\operatorname{max}(n \log n \log \log n, \frac{n^2}{s}))$.

Below we provide the algorithm for computing \Lis\ in pseudocode.
First we define a slightly modified version of \Ppass\ that will allow us to compute $L_{\leq x}$ and $L_{> x}$ for some $x$ by ignoring elements bigger than $x$ or elements smaller than or equal to $x$ respectively.
We will assume that if we are storing an element in working memory, then we also store its $\operatorname{pred}$.
\begin{algorithm}
    \caption{\Ppass\ -- the second pass through an array, computing $P_j$}
    \label{alg:ppas-mod}
    \begin{algorithmic}[1]
        \Require sequence $A$, space limit $s$, pile $P_i$, value $lb$, value $ub$, value of pile we are computing $j$

        \State set $h := 0$ 
        \State set $r := i$
        \State set $p_{i+1}, p_{i+2}, \ldots, p_{j}$ to $\infty$ 
        \State create an empty pile $P_j$
        \For{$a_t \in A$} 

            \If{$a_t$ is $P_i[h+1]$} 
                \State set $h := h + 1$ 
                \State continue
            \ElsIf {$a_t < P_i[h]$ or $a_t > p_j$ or $a_t > ub$ or $a_t \leq lb$} 
                \State continue
            \EndIf

            \If{$a_t > p_r$}
                \State set $r := r + 1$ and $k := r$
            \Else
                \State set $k$ to be the smallest such that $a_t < p_k$
            \EndIf
            
            \If{$t \leq \frac{|A|}{2}$}
                \State $\operatorname{pred}(a_t) := a_t$
            \ElsIf{$k = 1$}
                \State $\operatorname{pred}(a_t) := null$
            \Else 
                \State $\operatorname{pred}(a_t) := \operatorname{pred}(\operatorname{top}(P_{k-1}))$
            \EndIf

            \State set $p_k := a_t$

            \If{$k = j$}
                \State push $a_t$ onto $P_j$
            \EndIf
        \EndFor

    \Return $P_j$
    \end{algorithmic}
\end{algorithm}

Now, using this modified \Ppass\ procedure, we define the \LisLen\ procedure. It returns the $\operatorname{pred}$ value of an element from the last pile, which is the splitting value $m$ in our divide-and-conquer strategy.

\begin{algorithm}[H]
    \caption{\LisLen}
    \label{alg:lislen}
    \begin{algorithmic}[1]
        \Require sequence $A$, space limit $s$, values $lb, ub$
        \State set $i := 0$ and make an empty pile -- $P_0$
        \Repeat
            \State $j := \text{run first pass of \Ppass\ with } P_i, lb, ub$ \Comment{\textcolor{gray}{Find the next reachable pile $j$}}
            \State $P_j := \text{run second pass of \Ppass\ with } P_i, j, lb, ub$ 
            \State set $i := j$ and $P_i := P_j$ 
        \Until{$i$ is the length of \Lis}
        \State \Return $(i, \operatorname{pred}(\operatorname{top}(P_i)))$ \Comment{\textcolor{gray}{Return length and the element $m$}}
    \end{algorithmic}
\end{algorithm}

Finally, we present the main divide-and-conquer algorithm for extracting the \Lis. 
When the input sequence is small enough to fit in memory, we use the classic \Psort. 
Otherwise, we use \LisLen\ to find the element $m$, recursively retrieve the $L_{\leq m}$ for $A_\ell$ and $L_{>m}$ for $A_r$, and return their concatenation.
Initially we call the procedure with values $lb, ub$ set to $-\infty$ and $\infty$ respectively.

\begin{algorithm}[H]
    \caption{\Lis}
    \label{alg:lis-dc}
    \begin{algorithmic}[1]
        \Require sequence $A$, space limit $s$, values $lb, ub$
        \If{$|A| \leq s$} \Comment{\textcolor{gray}{Whole array fits in the working memory}}
            \State \Return classic \Lis\ of $A$ within $(lb, ub)$
        \EndIf
        
        \State $(k, m) := \LisLen(A, s, lb, ub)$ \Comment{\textcolor{gray}{Find length $k$ and splitting element $m$}}
        \State split $A$ into $A_\ell, A_r$ \Comment{\textcolor{gray}{Divide input into two halves}}
        
        \If{$m = null$} \Comment{\textcolor{gray}{The \Lis\ does not contain any elements from $A_\ell$}}
            \State \Return $\Lis(A_r, s, lb, ub)$
        \Else \Comment{\textcolor{gray}{Recursively solve for both halves using $m$ as a bound}}
            \State \Return $\Lis(A_\ell, s, lb, m) \parallel \Lis(A_r, s, m, ub)$ \Comment{\textcolor{gray}{$\parallel$ is concatenation}}
        \EndIf
    \end{algorithmic}
\end{algorithm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\section{Algorithm performance results}

\section{Summary}






\printbibliography

\end{document}